---
title: MATH70098 Ethics in Data Science and AI
subtitle: Exercise Sheet - Multi-objective optimisation and Pareto Fronts 
author: Zak Varty
#date: Autumn 2021
output: 
  pdf_document:
    pandoc_args: ["--lua-filter=color-text.lua"]
  html_document:
    pandoc_args: ["--lua-filter=color-text.lua"]
---
# Exercise sheet: Multi-objective optimisation and Pareto Fronts

The questions on this sheet are designed to let you test your own understanding of the course content on multi-objective optimisation and Pareto fronts. Some questions will test basic notions, while others will encourage you to think more deeply about some of the concepts introduced this week. 

--------------------------------------------------------------
```{r set-up, include=FALSE}

```
__Q1: Wiki contributions__ 

The Wikipedia article on [multi-objective optimisation](https://en.wikipedia.org/wiki/Multi-objective_optimization) provides a brief but comprehensive introduction to the topic. The ["Examples of applications" section](https://en.wikipedia.org/wiki/Multi-objective_optimization#Examples_of_applications) of this page lists application areas for multi-objective optimisation including economics, finance and optimal design. Along with each example application is a short one or two paragraph explanation of how multi-objective optimisation can be used in that area. Ethical AI is not currently on this list.

Write one to two paragraphs, in language suitable for this format, describing how multi-objective optimisation may be applied in the context of ethical AI. 

--------------------------------------------------------------

__Q2: Formal definitions of a Pareto-optimal and a dominated solution__

Formally define and explain in plain language what is meant by a dominated and a Pareto-optimal solution to the multi-objective optimisation problem: 

$$ \max_{x = (x_1,\ldots, x_n) \in X} g( f_1(x), f_2(x), \ldots f_k(x)) \text{ s.t. } x\in \mathcal{X}.$$


______________________________________________________________

__Q3: A concrete example__

Consider the size classifiers given in the table below. 

```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
library(gridExtra)
set.seed(1234)
example_df <- dplyr::tibble(
classifier = LETTERS[1:6],
k_anon = c(2,4,3,5,3,2), 
equalised_odds = c(F,T,T,F,T,T),
error_rate = c(0.11, 0.13, 0.05, 0.20, 0.13, 0.08),
false_positive_rate = round(runif(6) + c(0.4,-0.4,0,-0.45,0,0), 2)
)

grid.table(example_df)
```

 
(i) Consider the four pairs of objectives given below. Visualise the the bi-objective performance of the six classifiers for each of these pairs. Identify which classifiers are Pareto-optimal and which are dominated in each case.  

(a) false positive rate vs error rate,
(b) k-anonymity vs equalised odds,
(c) k-anonymity vs error rate,
(d) Equalised odds vs error rate. 

(ii) When considering all four objectives at once, which (if any) of the classifiers are Pareto optimal? 


-------------------------------------------------------------------------
__Q4: Scalarisation__  

(a) Define in your own words what it means to scalarise a multi-objective optimisation problem. 

(b) Linearisation and epsilon-constraint are two methods of linearisation. Explain in plain language how each method works. 

(c) Give linearised and epsilon-constrained formulations for the multi-objective optimisation in Question 2.










